---
title: "RExercise#2_dequito"
author: "Yanna Dawn Dequito"
date: "2026-09-30"
output: html_document
---

## Part A - Arithmetic Operations

```{r}
a <- 17
b <- 5

a + b
a - b
a * b
a / b

a %% b
a %/% b

2^10

```

**Question #1:** The normal division `17 / 5` gives the exact decimal answer, which is 3.4. The integer division `17 %/% 5` shows how many whole times 5 fits into 17, which is 3. The modulus `17 %% 5` gives the remainder of 2. Together, 5 times 3 plus the remainder 2 equals 17.

**Question #2:** I predicted `10 %% 3` would be 1 and `10 %/% 3` would be 3. I was correct because 3 goes into 10 exactly 3 times, leaving a remainder of 1.

## Part B - Basic Data Types

```{r}
x <- 42
y <- 42L
z <- "42"
w <- TRUE
v <- 3 + 2i

typeof(x)
typeof(y)
typeof(z)
typeof(w)
typeof(v)

class(x)
class(y)
class(z)
class(w)
class(v)

mode(x)
mode(y)
mode(z)
mode(w)
mode(v)

```

**Question #3:** `x <- 42` is stored as a "double" because R saves numbers as decimals by default. `y <- 42L` has an "L" at the end, which tells R to strictly save it as an "integer" instead.

**Question #4:** Yes, they can be different. For example, with `x <- 42`, `typeof(x)` says it is a "double", but `class(x)` and `mode(x)` say it is "numeric". The `typeof()` function gives the specific storage type, while the others are more general categories.

## Part C - Type Conversion

```{r}
num_string <- "123"
num_val <- as.numeric(num_string)
typeof(num_val)
class(num_val)
mode(num_val)

as.logical(0)
as.logical(1)

bad_string <- "hello123"
as.numeric(bad_string)

```

**Question #5:** R gives a warning and outputs `NA` instead of crashing so that the rest of our script can still run. The risk of ignoring these warnings is that the `NA` will secretly mess up our math calculations later on without us noticing.

**Question #6:** I think `as.logical(2)` and `as.logical(-1)` will both produce `TRUE`. After testing it, I see R follows a simple rule: the number 0 is `FALSE`, and every other number (positive or negative) becomes `TRUE`.

## Part D - Strings, paste(), and paste0()

```{r}
first_name <- "Yanna Dawn"
last_name <- "Dequito"

print(first_name)
print(last_name)
cat(first_name, last_name, "\n")

full_name <- paste(first_name, last_name)
full_name
paste0(first_name, last_name)

paste(first_name, last_name, sep = "-")

age <- 21
paste0(full_name, " is ", age, " years old.")

```

**Question #7:** The `paste()` function automatically adds a space between the words, while `paste0()` connects the words directly together with no spaces at all.

**Question #8:** It didn't cause an error because `paste0()` automatically converts the number `age` into a character string before putting it in the sentence, which is similar to using `as.character()`.

## Part E - Critical Thinking Challenge

```{r}
# Trying to run "5" + 5 gives a non-numeric argument error.
# Here is the fixed code using a conversion function:
as.numeric("5") + 5

```

**Question #9:** R requires us to convert it manually because it is a language made for strict math and statistics. It doesn't want to guess if we meant to add the numbers to get 10 or combine the text to get "55".

**Question #10:** The problem is probably that R read the numbers from the text file as character strings. To fix this, I would check the column type using `typeof()`. Then, I would use `as.numeric()` to convert the column into real numbers, check for any warnings, and finally try the `sum()` function again.

## Conclusion

In this lab, I learned how R categorizes data into different types and how to check them using functions like `typeof()` and `class()`. I also learned how to combine strings with `paste()` and `paste0()`, and how to manually convert data types so my math calculations don't throw errors. Understanding this is important so we can prepare and clean our data correctly before analyzing it.